---
title: Learning Python!
author: 'Noah Collaco '
date: '2020-05-08'
slug: learning-python
categories: []
tags: []
description: ''
---
It is possible to do most if not all the things that we did this semester in Rstudio in python. Below I will show functions that were done in my projects and try to recreate them in python. 
```{r setup, include=FALSE}
knitr::knit_engines$set(python= reticulate::eng_python)
library(reticulate)
py_install("matplotlib")

use_python("C:\\Users\\nkcol\\AppData\\Local\\r-miniconda\\envs\\r-reticulate\\python.exe")
```

```{R}
library(tidyverse)
library(readxl)
library(kableExtra)
library(GGally)
library(cluster)
library(readxl)
RankedMeasuredData <- read_excel("CopyOf2020 County Health Rankings Data - v1.xlsx", 
    sheet = "Ranked Measure Data")
AdditionalMeasureData <- read_excel("CopyOf2020 County Health Rankings Data - v1.xlsx", 
    sheet = "Additional Measure Data")
Education <- read_excel("CopyOfEducation.xls")
Education%>% filter(State=="TX")->TXEducation
RankedMeasuredData%>% filter(State=="Texas")-> RankedMeasuredData
AdditionalMeasureData%>% filter(State=="Texas")-> AdditionalMeasureData
TXEducation%>% rename(c("County"=`Area name`))->TXEducation
left_join(RankedMeasuredData,TXEducation, by="County")->partialdata
left_join(partialdata, AdditionalMeasureData, by="County")->fulldata

fulldata%>%select(County,`Life Expectancy`,`% Adults with Obesity`,`% Physically Inactive`,`% Smokers`,`% Food Insecure`,`% Uninsured`,`% Adults with Diabetes`,`Percent of adults with a bachelor's degree or higher, 2014-18` , `Median Household Income`, `% Insufficient Sleep`,`% Excessive Drinking`,`% Rural`)->project2
project2%>% na.omit()->project2
project2%>% rename(Lifeexpt=`Life Expectancy`)%>% rename(pctObese=`% Adults with Obesity`)%>% rename(pctInactive=`% Physically Inactive`)%>% rename(pctFDInsecure=`% Food Insecure`)%>% rename(pctSmokers=`% Smokers`)%>% rename(pctUninsured=`% Uninsured`)%>% rename(pctDiabetes=`% Adults with Diabetes`)%>% rename(pctBachelors=`Percent of adults with a bachelor's degree or higher, 2014-18`)%>% rename(medianincome=`Median Household Income`)%>% rename(pctinsufficientsleep=`% Insufficient Sleep`)%>% rename(pctRural=`% Rural`)%>% rename(pctexcessdrinking=`% Excessive Drinking`)->project2
project2%>% mutate(RurualorUrban=if_else(pctRural>50, "Mostly Rural", "Mostly Urban"))->project2


```
This is the data that was used in project two. I performed a linear regression an example to show how python and rstudio can both be used to perform statistical analysis. 

```{r}
fit1<-lm(Lifeexpt~pctObese ,data=project2)
summary(fit1)
```
```{python}
import pandas as pd
import numpy as np
from statsmodels.formula.api import ols
project2= r.project2
fit= ols("Lifeexpt~pctObese", data=project2).fit()
fit.summary()

```
This code shows linear regression of Life expectancy to percent Obesity in Texas counties. The linear regression was performed in both R and in python showing just one of the many similar ways R and python can be used. They both showed a significant negative relationship between obesity and life expectancy. Showing that each percentage increase of obesity results in a decrease of -.131 years. 
```{python}
import matplotlib.pyplot as plt
sns.boxplot(x="RuralorUrban", y="pctDiabetes", data=project2)
```