---
title: "Project 1"
author: "Noah Collaco nkc383"
date: "3/13/2020"
output:
  pdf_document: default
  html_document:
  df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, fig.align = "center", warning = F, message = F,
tidy=TRUE, tidy.opts=list(width.cutoff=60), R.options=list(max.print=100))
```

## R Markdown


```{R}
library(tidyverse)
library(readxl)
library(kableExtra)
library(GGally)
library(cluster)
Education <- read_excel("Education2.xls", 
    skip = 4)
RankedMeasureData<- read_excel("2019 County Health Rankings Texas Data - v1_0.xls", 
    sheet = "Ranked Measure Data")
AdditionalMeasureData<- read_excel("2019 County Health Rankings Texas Data - v1_0.xls", 
    sheet = "Additional Measure Data")


```
  This project uses data from the 2019 County Health Rankings for Texas and the USDA ERS Educational Attainment Data. These datasets are large so specific variables of County Name, % Obese, % Diabetic, Percent of a County with a Bachelors Degree, Median Household Income, and % Rural of a County were taken. 
```{r}

Education%>% filter(State=="TX")->TXEducation
TXEducation%>% rename(c("County"=`Area name`))->TXEducation
left_join(RankedMeasureData, TXEducation, by="County")->partialdata
left_join(partialdata, AdditionalMeasureData, by="County")->fulldata
head(fulldata)
fulldata%>%select(County, `% Obese`,`% Diabetic`,`Percent of adults with a bachelor's degree or higher, 2014-18` , `Household Income`, `% Rural`)->project1

project1%>% slice(2:255)->project1
project1%>% rename(PercentObese=`% Obese`)%>% rename(PercentDiabetic=`% Diabetic`)%>%
  rename(PercentBachelors=`Percent of adults with a bachelor's degree or higher, 2014-18`)%>%
  rename(Medianincome=`Household Income`) %>% rename(pctRural=`% Rural`)->project1
glimpse(project1)
```
To create the data set that was used for project one a left join was first performed with 2019 County Health Rankings Ranked Measure Data with data from the USDA Economic Research Service about educational attainment, which had already been filtered for TX counties only. This partial dataset was then joined, using a left join, to 2019 County Health Rankings Additional Measured data to in order to get median household income data.The 5 variables that were selected for the project were % Obesse, % Inactive, % Bachelors Degree, Median Household Income, and % Rural. Left Joins were used to keep all columns for both dataframs but only for the counties that matched between each dataframe. 


```{R}
project1%>% arrange(-Medianincome)%>% glimpse()
project1%>% mutate(`Rural/Urban`=if_else(pctRural>50, "Mostly Rural", "Mostly Urban"))->project1
glimpse(project1)

```

Summary Statistics 
Pivot wider and longer are used here to create organized tables of summary statistics. Pivot longer takes the names and values of the summary statistics into two columns called variable and stat. Pivot wider takes the names from variable and the values fro value to create a new column. 
```{R}
project1 %>% summarise_at(2:5,funs(mean=mean, sd=sd, var=var, max=max, min=min))
project1 %>% summarise_at(2:5,funs(mean=mean, sd=sd, var=var, max=max, min=min))%>% pivot_longer(contains("_"))%>%separate(name,sep="_",into=c("variable","Stat"))%>% pivot_wider(names_from = "variable",values_from="value")%>%arrange(Stat)%>% kable()
                         
project1%>% group_by(`Rural/Urban`)%>%summarise_at(2:5,funs("mean"=mean, "sd"=sd, "var"=var, "max"=max, "min"=min))%>% pivot_longer(contains("_"))%>%separate(name,sep="_",into=c("variable","Stat"))%>% pivot_wider(names_from = "variable",values_from="value")%>%arrange(Stat)%>% kable()

project1%>% select_if(is.numeric)%>%cor()%>% kable()


```

```{R}
project1%>%select_if(is.numeric)%>%cor() %>%as.data.frame%>%
rownames_to_column%>% pivot_longer(-1,names_to="name",values_to="correlation")->tidycor
tidycor%>% ggplot(aes(rowname,name,fill=correlation))+geom_tile()+scale_fill_gradient2(low="red",high="blue")+ geom_text(aes(label=round(correlation,2)),color = "black", size = 4)+ theme(axis.text.x = element_text(angle = 90, hjust = 1))+ coord_fixed()+ ggtitle("Correlation Matrix for TX County Data")
```
  
This table shows the correlations between each variable within the project 1 dataset. In blue/purple are the positive correlations between variables. The squares in red/pink show negative correlations. Percent Bachelors positively correlates with Median Household income, and negatively correlates with percent rural, percent diabetic, and percent obese. Percent rural positively correlates with percent diabetic, and negatively correlates with median income, percent bachelors, and percent obesity. Percent obese positively correlates with percent diabetic, but negatively correlates with median income, percent rural, and percent bachelors. These correlations helped to guide what visualizations that were chosen. 

```{R}
project1 %>% ggplot()+geom_point( aes(x=Medianincome , y=PercentBachelors,color=`Rural/Urban`, ))+scale_x_continuous()+geom_smooth(method='lm', se=FALSE, aes(x=Medianincome , y=PercentBachelors,color=`Rural/Urban`)) +ggtitle("Percent Bachelors vs Median Income in Texas Counties") +scale_x_continuous(labels=scales::dollar)+xlab("Median Household Income")+ylab("% of County with Bachelors Degree")+scale_y_continuous(breaks=seq(0, 50, 10))+scale_color_brewer(palette = "Set1")

```
  This graph displays a scatterplot of Median Household Incomes of Texas counties with Percent of a county holding a bachelors degree. There is a positive relationship between having a higher percentage of people with a bachelors degree in a county and a higher median income as seen by the trendline. The "Mostly Urban" counties also have a trendline with a more positive slope compared to the "Mostly Rural" Counties. This indicates that the positive relationship between Percent bachelors and Median household income is stronger in Urban Counties than Rural Counties. 

```{R}
project1 %>% ggplot()+geom_smooth(method='lm', se=FALSE, aes(x=PercentObese , y=PercentBachelors,color=`Rural/Urban`))+geom_point( aes(x=PercentObese , y=PercentBachelors ,color=`Rural/Urban`, size=Medianincome))+scale_x_continuous() +ggtitle("Percent Bachelors vs Obesity Rate in Texas Counties") +xlab("Obesity Rate (%)")+ylab("% of County with Bachelors Degree")+scale_color_brewer(palette = "Paired")

```
  This graph shows the relationship between Obesity Rate, % of a county with a bachelors degree, and median household income. There is a clear negative relationship between obesity rate and Percent bachelors degree. As % Bachelors decreases obesity rates increase as seen by the trendlines. The size of the points indicate miedian income as seen through the bigger points having higher educational attainment. There is also a difference btween Rural/Urban counties as Urban counties have a stronger relationship between obeisty and bachelors degree than rural counties as seen throught the differing slopes. 

```{R}

project1%>% ggplot(aes(x=`Rural/Urban`, y=PercentDiabetic,fill=`Rural/Urban`))+geom_bar(stat = "summary",fun.y="mean", width = .7)+geom_errorbar(stat="summary", width=.5)+theme(legend.position = "none") +ggtitle("Mean Diabetic Percentage in Rural vs Urban Counties") +scale_fill_brewer(palette="Set1")

```
  This graph shows a clear difference between Mostly Rural and Mostly Urban counties in the rates of diabetes. Urban counties on average have lower rates of Diabetes than Rural counties. 

```{R}

clust_dat<-project1%>% select(2:6)
sil_width<-vector()
for(i in 2:10){    
  pam_fit <- pam(clust_dat, k = i)    
  sil_width[i] <- pam_fit$silinfo$avg.width  
}
ggplot()+geom_line(aes(x=1:10,y=sil_width))+scale_x_continuous(name="k",breaks=1:10) +ggtitle("Silhouette Width vs Number of Clusters")
```
  This is a graph of silhouette width for the numeric data for project one. This graph shows that 2 clusters would be the optimal number of clusters as it has the highest silhouette width.
```{R}
pamproject1<-clust_dat%>% pam(k=2)
pamproject1
```
```{R}
project1%>% count(`Rural/Urban`)
pamclust<-clust_dat %>% mutate(cluster=as.factor(pamproject1$clustering))
confmat<-pamclust%>%mutate(RuralUrban=project1$`Rural/Urban`)%>%group_by(RuralUrban)%>%count(cluster)%>% pivot_wider(names_from="cluster",values_from="n",values_fill = list('n'=0))
confmat

```
  Based on the clustering data it does not seem that the counties are clustering based on their rural/urban distinction. There must be a different factor that accounts for the clustering data seen. The first table shows the counts of Mostly Rural and Mostly Urban Counties.
```{R}
plot(pamproject1)
```
  This graph indicated the average silhouette width is .58 which can be interpreted as a resonable structure has been found among the clusters. 
```{R}
project1%>%mutate(cluster=pamproject1$clustering)%>%rename_all(function(x)sub("_", ".", x))%>%select_if(is.numeric)%>% group_by(cluster)%>%mutate(n=n())%>%group_by(cluster,n)%>% summarize_all( .funs = list("mean"=mean,"median"=median,"sd"=sd),na.rm=T)%>% pivot_longer(contains("_"))%>%separate(name,sep="_",into=c("variable","stat"))%>% pivot_wider(names_from = "variable",values_from="value")%>%arrange(stat)%>% kable()
```
The table above displays summary statistics for each cluster. 
```{R}
clust_dat %>% mutate(cluster=as.factor(pamproject1$clustering))%>%ggpairs(columns=1:5, aes(color=cluster))

```
This GGpairs plot shows the two different clusters as red and blue. From this visualization median income is the main differentiating factor between the two clusters. 
```{R}
pamclust<-clust_dat %>% mutate(cluster=as.factor(pamproject1$clustering))
pamclust %>% mutate(`Rural/Urban`=project1$`Rural/Urban`)%>%ggplot(aes(PercentObese,PercentBachelors,color=`Rural/Urban`, shape=cluster)) + geom_point() +scale_color_brewer(palette = "Set1")+ ggtitle("Percent Bachelors vs Obesity with Cluster Data")+ylab("% of County with Bachelors Degree")+xlab("Obesity Rate (%)")
```
  This visualization does not show any clear distinction between clusters. This means that the clusters are more likely based on variable(s) that are not % Bachelors, % Obese, or Rural/Urban.
```{R}
pamclust %>% mutate(`Rural/Urban`=project1$`Rural/Urban`)%>%ggplot(aes(PercentBachelors,Medianincome,color=`Rural/Urban`, shape=cluster)) + geom_point() +scale_color_brewer(palette = "Set1")+ ggtitle("Percent Bachelors vs Median Income with Cluster Data")+scale_y_continuous(breaks=seq(20000, 100000, 10000),labels=scales::dollar)+geom_abline(slope=1, intercept = 50000)+xlab("% of County with Bachelors Degree")+ylab("Median Household Income")
```

  This second visualization captures the distinction between the clusters. The clusters seem to be determined by Median Household Income with the cutoff between clusters occuring around $50,000. Median Income seems to be the main differentiating factor between these two clusters. 



